Structuring Theories with Implicit Morphisms

### 2019-02-21 intro.tex

\paragraph{Contribution}
At the highest level, our contribution is the observation that implicit morphisms form a sweet spot of a very simple language feature that has substantial practical uses.
We recommend using implicit morphisms in all theory morphism--based formalisms.
More concretely, we present a formal system for developing structured theories with implicit morphisms.
Our starting point is the \mmt language \cite{RK:mmt:10}, which already provides a very general setting for defining and working with theories and morphisms.
\mmt is logic-independent, i.e., allows embedding a large variety of declarative languages (logics, type-theories, etc.).
Therefore, all our results can be directly applied to any language $L$ represented in \mmt or easily transferred to dedicated implementations of $L$.

>>>>

\paragraph{Contribution}
At the highest level, our contribution is the observation that implicit morphisms form a sweet spot of a very simple language feature that has substantial practical uses.
We recommend using implicit morphisms in all theory morphism--based formalisms.
More concretely, we present a formal system for developing structured theories with implicit morphisms based on the \mmt language \cite{RK:mmt:10}.
We choose \mmt because it provides a simple framework for working with theories and morphisms while remaining logic-independent: \mmt allows embedding a large variety of declarative languages $L$ (logics, type-theories, etc.) so that our results immediately yield an implementation of implicit morphisms for any such $L$.
We also expect but do not describe here that other morphism-based systems can be easily extended to allow for implicit morphisms.

### 2019-02-21 paper.tex

\begin{abstract}
We introduce \emph{implicit} morphisms as a concept in formal systems based on theories and theory morphisms.
The idea is that there may be at most one implicit morphism from a theory $S$ to a theory $T$, and if $S$-expressions are used in $T$ their semantics is obtained by automatically inserting the implicit morphism.
The practical appeal of implicit morphisms is that they hit the sweet-spot of being extremely simple to understand and implement while significantly helping with structuring large collections of theories.
Concrete applications include elegantly identifying isomorphic theories and extending theories with definitions and theorems as well as efficiently building and maintaining large, fine-granular, and heterogeneous hierarchies of theories.
\end{abstract}

@@

\bibliographystyle{alpha}
\bibliography{macros/pub_rabe,macros/rabe,macros/systems}

>>>>

\begin{abstract}
We introduce \emph{implicit} morphisms as a concept in formal systems based on theories and theory morphisms.
The idea is that there may be at most one implicit morphism from a theory $S$ to a theory $T$, and if $S$-expressions are used in $T$ their semantics is obtained by automatically inserting the implicit morphism.
The practical appeal of implicit morphisms is that they hit the sweet-spot of being extremely simple to understand and implement while significantly helping with structuring large collections of theories.

Concrete applications include elegantly identifying isomorphic theories and extending theories with definitions and theorems as well as efficiently building and maintaining large, fine-granular, and heterogeneous hierarchies of theories.
Our results are formulated and implemented in the \mmt language and system, and we expect they can be transfered to other morphism-based formalisms relatively easily.
\end{abstract}

@@

\bibliographystyle{alpha}
\bibliography{macros/pub_rabe,macros/rabe,macros/systems,macros/institutions}
